---
title: "Exercises Laboratory Session 03"
author: "Nicola Zomer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
---

```{r setup-chunk}
knitr::opts_chunk$set(dev = "ragg_png")
options(digits=5) # set number of digits equal to 5

```

# Packages and functions

```{r, message=FALSE}

# tidyverse
library(tidyverse)

# others
library(gridExtra)
library(kableExtra)
library(glue)
library(highcharter)
# library(scales)

```

Define a function for printing data frames with kable:

```{r}

mykable <- function(data) {
  knitr::kable(data) %>% 
    kable_styling()
}

```


# Exercise 1


### a) plot the function, given the interval (a, b)

```{r}

f <- function(x, a, b, c){
  if (!(c>=a & c<=b)) stop('c is not between a and b')
  
  ifelse(
    a<=x & x<c, 
    2*(x-a)/((b-a)*(c-a)), 
    ifelse(
      c<=x & x<=b, 
      2*(b-x)/((b-a)*(b-c)), 
      0
    )
  )
}

a = -2
b = 2
c = 0

x_plot = seq(-3, 3, by=0.2)
ggplot()+
  geom_line(aes(x=x_plot, y=f(x_plot, a, b, c)), linetype='dashed')+
  geom_point(aes(x=x_plot, y=f(x_plot, a, b, c)))+
  labs(
    title='Triangular distribution, with a=-2, b=2, c=0', 
    x='X', 
    y='f(X)'
  )+
  scale_x_continuous(breaks=seq(-3, 3, by=0.5))+
  theme_bw()




```

### b) and write an algorithm to generate random numbers from the triangular distribution
### c) generate 104 random number from the distribution, show them in an histogram and superimpose the analytical curve

# Exercise 2 - Markov’s inequality
Plot G(k) and the Markov’s upper bound for
a) the exponential, $Exp(\lambda = 1)$, distribution function
b) the uniform, $U(3, 5)$, distribution function
c) the binomial, $Bin(n = 1, p = 1=2)$, distribution function
d) a Poisson, $Pois(\lambda = 1=2)$, distribution function

### a) 
### b)
### c)
### d)

# Exercise 3 - Chebyshev’s inequality
# Exercise 4 - Six Boxes Toy Model : inference
# Exercise 5 - Six Boxes Toy Model : simulation


