---
title: "Exercises Laboratory Session 02"
author: "Nicola Zomer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
---

```{r setup-chunk}
knitr::opts_chunk$set(dev = "ragg_png")

```


```{r}
# tidyverse
library(ggplot2)

# others
library(kableExtra)

```

Define a function for printing data frames with kable:

```{r}

mykable <- function(data) {
  knitr::kable(data) %>% 
    kable_styling()
}

```


# Exercise 1
```{r}

x <- c(15.58, 15.9, 16, 16.1, 16.2)
p1 <- c(0.15, 0.21, 0.35, 0.15, 0.14)
p2 <- c(0.14, 0.05, 0.64, 0.08, 0.09)

exp_value_1 <- sum(x*p1)
exp_value_2 <- sum(x*p2)

var_1 <- sum(p1*(x-exp_value_1)**2)
var_2 <- sum(p2*(x-exp_value_2)**2)

cat(
  'E[X] method 1:', round(exp_value_1, 3), 
  '\nE[X] method 2:', round(exp_value_2, 3), 
  '\nVar(X) method 1:', round(var_1, 6), 
  '\nVar(X) method 2:', round(var_2, 6)
  )


```

# Exercise 2
The waiting time, in minutes, at the doctor's is about 30 minutes, and the distribution follows an exponential pdf with rate 1/30.

## A) Simulate the waiting time for 50 people at the doctor's office and plot the relative histogram

```{r}

set.seed(1) 
people_wt <- rexp(50, 1/30)

ggplot() + 
  geom_histogram(aes(people_wt),  bins = 10, fill='dodgerblue3', colour="black") +
  labs(title="Simulation of the waiting time for 50 people", 
       x='People Waiting Time',
       y='Count')+
  scale_x_continuous(
    breaks = seq(0, 160, 20),
  ) + 
  theme_bw()

```

## B) What is the probability that a person will wait for less than 10 minutes?

```{r}

```


## C) Evaluate the average waiting time from the simulated data and compare it with the expected value 
Calculate the expected value from theory and by manipulating the probability distributions using R.

```{r}

```


## D) What is the probability for waiting more than one hour before being received?

```{r}

```



# Exercise 3

# Exercise 4

# Exercise 5

# Exercise 6


