---
title: "Exercises Laboratory Session 03"
author: "Nicola Zomer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
---

```{r setup-chunk}
knitr::opts_chunk$set(dev = "ragg_png")
options(digits=5) # set number of digits equal to 5

```

# Packages and functions

```{r, message=FALSE}

# tidyverse
library(tidyverse)

# others
library(gridExtra)
library(kableExtra)
library(glue)
library(highcharter)
library(latex2exp)

options(highcharter.theme = hc_theme_elementary())

```


# Exercise 1 - Community Mobility Open Data
Community Mobility Reports have been created with the aim to provide insights into what has changed in response to policies aimed at combating COVID-19. Data can be found at https://www.google.com/covid19/mobility/.

Download and analyze the following data sets:

* https://www.gstatic.com/covid19/mobility/Global_Mobility_Report.csv 
* https://www.gstatic.com/covid19/mobility/Region_Mobility_Report_CSVs.zip

Select a couple of European countries of your choice and analyze the trends in the variables over time: produce a plot of the data by averaging the observable over a period of one week (hint: convert the data field to `lubridate::week`) and one month and quantify the impact of COVID-19 restrictions on mobility situations.

```{r}

```


# Exercise 2 - Random number generators
One of the first random number generator was proposed by von Neumann, the so-called _middle square_ algorithm. 

Write R code to implement this type of generator and, given a fixed digit number input, square it an remove the leading and trailing digits, in order to return a number with the same number of digits as the original number.

```{r}

middle_square <- function(x.start=5772156649, n){
  input.digits <- length(unlist(strsplit(as.character(x.start),"")))
  
  # print(paste('Number of digits of the input number:', input.digits))
  if (x.start%%1 != 0) stop('The starting number must be integer')
  
  x.squared.char <- unlist(strsplit(as.character(x.start^2),""))  
  half.output.digits <- length(x.squared.char)%/%2
  
  if (input.digits%%2 ==0){ # even number of digits
    x.i <- x.squared.char[(half.output.digits-(input.digits/2-1)):(half.output.digits+input.digits/2)]
  } else{ # odd number of digits
    x.i <- x.squared.char[(half.output.digits-(input.digits%/%2)):(half.output.digits+input.digits%/%2)]
  }
  
  # check if the first digit is zero: in this case shift to the left
  i<-1
  while (x.i[1]=="0"){
    print(x.i)
    if (input.digits%%2 ==0){ # even number of digits
      x.i <- x.squared.char[(half.output.digits-(input.digits/2-1)-i):(half.output.digits+input.digits/2-i)]
    } else{ # odd number of digits
      x.i <- x.squared.char[(half.output.digits-(input.digits%/%2)-i):(half.output.digits+input.digits%/%2-i)]
    }
    i<-i+1
  }
  return(as.numeric(paste(x.i, collapse="")))
}

```

Generate 10 numbers using the _middle square_ algorithm:
```{r}

numbers <- numeric(11)
numbers[1] <- 5772156649

for (i in 1:10){
  numbers[i+1] <- middle_square(numbers[i])
}

data.frame(Iteration=0:10, Generated.Number=numbers) %>%
  kable() %>%
  kable_styling(full_width=FALSE)

```

# Exercise 3 - Bayesian Inference

A publishing company has recently launched a new journal. In order to determine how effective it is in reaching its possible audience, a market survey company selects a random sample of people from a possible target audience and interviews them. Out of 150 interviewed people, 29 have read the last issue of the journal.

### a) What kind of distribution would you assume for y, the number of people that have seen the last issue of the journal?

```{r}

```

### b) Assuming a uniform prior, what is the posterior distribution for y ?

```{r}

```

### c) Plot both posterior and likelihood ditributions functions

```{r}

```


# Exercise 4 - Bayesian Inference
A coin is flipped n = 30 times with the following outcomes:

T, T, T, T, T, H, T, T, H, H, T, T, H, H, H, T, H, T, H, T, H, H, T, H, T, H, T, H, H, H

### a) Assuming a flat prior, and a beta prior, plot the likelihood, prior and posterior distributions for the data set.

```{r}

```

### b) Evaluate the most probable value for the coin probability p and, integrating the posterior probability distribution, give an estimate for a 95% credibility interval.

```{r}

```

### c) Repeat the same analysis assuming a sequential analysis of the data. Show how the most probable value and the credibility interval change as a function of the number of coin tosses (i.e. from 1 to 30).

```{r}

```

### d) Do you get a different result, by analyzing the data sequentially with respect to a one-step analysis (i.e. considering all the data as a whole)?

```{r}

```













