---
title: "Exercises Laboratory Session 03"
author: "Nicola Zomer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
---

```{r setup-chunk}
knitr::opts_chunk$set(dev = "ragg_png")
options(digits=5) # set number of digits equal to 5

```

# Packages and functions

```{r, message=FALSE}

# tidyverse
library(tidyverse)

# others
library(tictoc)
library(gridExtra)
library(kableExtra)
library(glue)
library(highcharter)

```

Define a function for printing data frames with kable:

```{r}

mykable <- function(data) {
  knitr::kable(data) %>% 
    kable_styling()
}

```


# Exercise 1
Consider the following triangular distribution
$$
f(x)=
\begin{cases}
  \frac{2(x-a)}{(b-a)(c-a)} & a\leq x < c \\
  \frac{2(b-x)}{(b-a)(b-c)} & c\leq x < b \\
  0 & \text{otherwise}
\end{cases}
$$
where $c\in[a, b]$.

### a) Plot the function, given the interval (a, b)

```{r}

d_triang <- function(x, params){
  a = params[1]
  b = params[2]
  c = params[3]
  
  if (!(c>=a & c<=b)) stop('c is not between a and b')
  
  ifelse(
    a<=x & x<c, 
    2*(x-a)/((b-a)*(c-a)), 
    ifelse(
      c<=x & x<=b, 
      2*(b-x)/((b-a)*(b-c)), 
      0
    )
  )
}

a = -2
b = 2
c = 0

x_plot = seq(-3, 3, by=0.1)

# generic plot of a function
myplot <- function(x_values, funct, args, title='My function', xlab = 'x', ylab='f(x)'){
  ggplot()+
    geom_line(aes(x=x_values, y=funct(x_values, args)))+
    labs(
      title=title, 
      x=xlab, 
      y=ylab
    )+
    scale_x_continuous(breaks=seq(x_values[1], tail(x_values, 1), by=0.5))+
    theme_bw()
  
}

myplot(x_plot, d_triang, c(a, b, c), title='PDF of the Triangular distribution, with a=-2, b=2, c=0', xlab='X', ylab='f(X)')


```

### b) Write an algorithm to generate random numbers from the triangular distribution


#### Inverse transform method
```{r}

# Cumulative distribution
p_triang <- function(x, params){
  integrals = numeric(length(t))
  for (i in seq_along(x)){
    integrals[i] <- integrate(d_triang, lower=-Inf, upper = x[i], params)$value
  }
  return(integrals)
}

myplot(x_plot, p_triang, c(a, b, c), title='CDF of the Triangular distribution, with a=-2, b=2, c=0', xlab='X', ylab='F(X)')

```


```{r}

# Inverse functions;
# bounds set such that it searches for a solution in [-2, 2], the only interval in which the cdf is invertible
inverse <- function(y, params){
    uniroot(function(x){p_triang(x, params) - y}, lower = -2, upper = 2)$root
}

# Quantile function
q_triang <- function(y, params, lower.value=-2, upper.value=2){
  output <- numeric(length(y))
  for (i in seq_along(y)){
    if (y[i]==1) output[i]<-upper.value       # boundary value
    else if (y[i]==0) output[i]<-lower.value  # boundary value
    else output[i] <- inverse(y[i], params)
  }
  return(output)
}

# Generate random numbers 
r_triang_1 <- function(n, params, seed=1){
  set.seed(seed)
  q_triang(runif(n), params)
}

```

#### Acceptance/rejection method
In order to use the acceptance/rejection method, we must find a finite interval $[a', b']$ in which the pdf is defined and a value $M\in\mathbb{R}$ such that $\forall x \in [a',b'] \; f(x)<M$. 

My choices for the interval and $M$ are:
- $[a', b'] = [a, b] = [-2, 2]$
- $M=\max_{x\in[a, b]} f(x) = f(x=0) = 0.5$ 

I select  interval $[a, b] = [-2, 2]$ and $M$ to be the maximum value of the pdf, i.e. $f(x=0)$

```{r}
d_triang(0, c(a, b, c))
```


```{r}

r_triang_2 <- function(n, f, params, seed=1, a_=a, b_=b, M=0.5){
  
  set.seed(seed)
  y = numeric(n)

  i=1
  while (i <= n){
    
    u_1 <- runif(1, a_, b_)
    u_2 <- runif(1, 0, 1)
    x_1 <- a + (b-a)*u_1
    
    if (u_2*M < f(u_1, params)){
      y[i] <- u_1
      i=i+1
    }
  }
  return(y)
}



```


### c) Generate $10^4$ random number from the distribution, show them in an histogram and superimpose the analytical curve
#### Inverse transform method

```{r}

tic("Generate 10^4 random numbers using the inverse transform method")
generated_points <- r_triang_1(10000, c(a, b, c))
toc()

```


```{r, warning=FALSE}

gg_gen <- function(method){
  ggplot() + 
    geom_histogram(aes(x=generated_points, y=..density..),  binwidth = 0.1, center = 0.1, fill='lightblue', colour="black") +
    stat_function(fun=function(x)d_triang(x, c(a, b, c)), color='firebrick', size=0.6) +
    labs(title=paste('Sampling results using the', method, 'method'), 
        x='X',
        y='Count (normalized)')+
    scale_x_continuous(breaks=seq(-3, 3, by=0.5), limits=c(-3,3)) +
    theme_bw()
}

gg_gen('inverse transform')

```


#### Acceptance/rejection method
```{r, warning=FALSE}

tic("Generate 10^4 random numbers using the acceptance/rejection method")
generated_points <- r_triang_2(10000, d_triang, c(a, b, c))
toc()

gg_gen('acceptance/rejection')

```


# Exercise 2 - Markov’s inequality
Markov’s inequality represents an upper bound to probability distributions. Having defined a function $G(k) = 1-F(k) \equiv P(X \geq k)$, plot $G(k)$ and the Markov’s upper bound for

* **a)** the exponential, $Exp(\lambda = 1)$, distribution function
* **b)** the uniform, $\mathcal{U}(3, 5)$, distribution function
* **c)** the binomial, $Bin(n = 1, p = 1/2)$, distribution function
* **d)** a Poisson, $Pois(\lambda = 1/2)$, distribution function

### a) Exponential distribution function
### b) Uniform distribution function
### c) Binomial distribution function 
### d) Poisson distribution function

# Exercise 3 - Chebyshev’s inequality
Use R to show, with a plot, that Chebyshev’s inequality is is an upper bound to the following distributions:

* **a)** a normal distribution, $\mathcal{N}(\mu=3, \sigma=5)$
* **b)** an exponential distribution $Exp(\lambda = 1)$
* **c)** a uniform distribution, $\mathcal{U}(1-\sqrt{2}, 1+\sqrt{2})$
* **d)** a Poisson, $Pois(\lambda = 1/3)$, distribution function

### a) Normal distribution 
### b) Exponential distribution 
### c) Uniform distribution  
### d) Poisson distribution 

# Exercise 4 - Six Boxes Toy Model : inference
The six boxes toy model is described in [[1]](https://arxiv.org/pdf/1612.05292.pdf).

Write a program in R that:

1. allows the user to insert the color of a randomly extracted box and
2. prints on the standard output the probability of selecting each box
3. plots the probability for each box as a function of the extraction step

# Exercise 5 - Six Boxes Toy Model : simulation
Consider again the six boxes toy model of the previous exercise and write a simulation program that:

1. selects a random box
2. makes random sampling from the box
3. prints on the standard output the probability of selecting each box
4. plots the probability for each box as a function of the number of trial


# Bibliography
<dl>
  <dt>[1]</dt>
  <dd>
    G. D’Agostini, Probability, propensity and probabilities of propensities (and of probabilities), https://arxiv.org/pdf/1612.05292.pdf
  </dd>
  <dt>[2]</dt>
  <dd>
    G. D’Agostini, More lessons form the six box toy experiment, https://arxiv.org/pdf/1701.01143.pdf
  </dd>
</dl>